---
title: "ML Wisconsin Breast Cancer Dataset"
author: "Michael Harrison"
date: "May 1, 2017"
output: html_document
---


# 1. Prepare Problem
# a) Load packages
```{r}
library(mlbench)
library(caret)
```

# b) Load dataset
```{r}
data("BreastCancer")
```

# c) Split-out validation dataset
```{r}
inTrain <- createDataPartition(BreastCancer$Class, p = 0.8, list=FALSE)
training <- BreastCancer[inTrain,]
testing <- BreastCancer[-inTrain,]
```

# 2. Summarize Data
# a) Descriptive statistics
- Training set dimensions
```{r}
dim(training)
```
- Head of dataset
```{r}
head(training, 10)
```

- Checking classes of attributes
```{r}
sapply(training, class)
```
- 
```{r}
#Remove ID attribute
training <- training[,-1]
#Convert input values to numeric
for(i in 1:9){
        training[,i] <- as.numeric(as.character(training[,i]))
}
```

- Summary of data
```{r}
summary(training)
```

- Class distribution
```{r}

```

# b) Data visualizations

# 3. Prepare Data
# a) Data Cleaning
# b) Feature Selection
# c) Data Transforms

# 4. Evaluate Algorithms
# a) Test options and evaluation metric
# b) Spot-Check Algorithms
# c) Compare Algorithms

# 5. Improve Accuracy
# a) Algorithm Tuning
# b) Ensembles

# 6. Finalize Model
# a) Predictions on validation dataset
# b) Create standalone model on entire training dataset
# c) Save model for later use